\sysfull

### 2018-08-03 paper/asplos19/background.tex
partial commit

Energy harvesting devices operate using energy extracted from sources such as
radio frequency transmissions and solar energy. These devices elide tethered
power or a battery, instead collecting energy into a capacitor, operating when
sufficient energy accumulates, and upon depleting the energy, turning off and
recharging.
%
%Batteryless operation has a number of important advantages, making intermittent computing an important research domain. Supplying power to billions~\cite{gartner_iot} of embedded computers using batteries is not sustainable. The European Commission estimates that more than 160 kilotons of consumer batteries enter the European Union annually~\cite{eu_batteries_2016}. Batteries are an environmental risk, are fragile, are limited in their number of charge/discharge cycles, and may require costly physical maintenance that is difficult or impossible deployed (e.g. in space~\cite{kicksat}). By contrast, super-capacitors are durable, promising millions of charge/discharge cycles~\cite[Sec. I]{ongaro_pwre_2012}. The main limitation in moving to capacitor-based energy storage is that capacitor energy density---and consequently operating discharge time---is orders of magnitude less than a battery. 
%
%Given current technology development, battery-less systems are best suited for very long-term sensing and monitoring where access to recharge is either prohibitive or impossible. These include battery-less image capture and processing~\cite{naderiparizi_rfid_2015}, animal monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015} and digestible~\cite{nadeau_naturebio_2017} sensors.
%
There are several energy harvesting battery-less hardware platforms. For
instance, computational RFIDs---open-source TI MSP430-based~\cite{wolverine}
WISP~\cite{wisp5} (with its variants such as
WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or
NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones
such as~\cite{medusa_farsens_2017}. Other platforms include ambient backscatter
tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less
phone~\cite{talla_imwut_2017}.

%In all of the above, the main source of energy harvested is the electromagnetic radiation in the radio frequency range (ambient transmitters such as high power TV transmitters~\cite{liu_sigcomm_2013} or dedicated RFID antenna~\cite{wisp5,moo,talla_imwut_2017,medusa_farsens_2017,holleman_biocas_2008,naderiparizi_rfid_2015}). Naturally, other forms of energy harvesting sources exist, including temperature gradient, (micro-)motions, light/sun radiation, vibrations, and body fluid flow (blood, gastric acid). Several recent surveys discussing energy harvesting, low-power, embedded systems and intermittent computing at a high level~\cite{paradiso_pvc_2005,soyata_csm_2016,prasad_comst_2014,ku_cst_2016,lucia_snapl_2017}.

@@

\begin{figure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-example.pdf}
        \caption{WAR-affected code}
        \label{fig:war-example}
    \end{subfigure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
        \caption{Incorrect execution}
        \label{fig:war-execution}
    \end{subfigure}
    \caption{WAR dependence example. \texttt{NV} marks non-volatile variables,
    while \texttt{checkp()} is a checkpoint of volatile state.}
    \label{fig:war}
\end{figure}

Software running on an energy-harvesting device executes {\em intermittently}
because energy sources are not always available to harvest and buffer
sufficient operational energy. An intermittent execution is composed of
operating periods interspersed with power
failures~\cite{dino,chain,alpaca,ratchet}. The frequency of failures depends on
the size of the device's energy storage buffer: a larger buffer allows longer
operating periods. 
%Energy-harvesters provide input power orders of magnitude less than operating power, making recharging negligible during operation.
%Intermittent execution is different from continuous execution. 
A power failure clears volatile state (e.g., registers and SRAM) while
non-volatile memory (e.g., FRAM) persists. Upon a power failure, control flows
to a prior point in the execution: by default, to the beginning of {\tt
main()}. Early intermittent systems preserved progress by periodically
checkpointing volatile execution context to non-volatile
memory~\cite{mementos}, sometimes using hardware
support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}.

%\begin{figure}
%	\centering
%	\subfloat[Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration]{\includegraphics[width=\columnwidth]{figures/crc_example}\label{fig:crc_example}}\\
%	\subfloat[Consecutive execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)]{\includegraphics[width=\columnwidth]{figures/crc_example_war}\label{fig:crc_example_war}}
%	\caption{Code example demonstrating effect of write after read on volatile memory checkpointing.}
%	\label{fig:code_demo_incosistency}
%\end{figure}

Checkpointing volatile state alone does not ensure data consistency when the
system can directly manipulate non-volatile memory~\cite{mspcdino}.
Precisely, data can get inconsistent when code includes a
\emph{write-after-read} (WAR) dependence between operations that manipulate
non-volatile memory, because it allows {\em writes} from a failed attempt to
persist across failure and be read by the {\em
reads}~\cite{ratchet,dino,alpaca}.
Figure~\ref{fig:war} illustrates how state can become
inconsistent in an intermittent execution using a simple example of an average
operation over an array of integers.
The variable \texttt{sum} introduces the WAR, being read and written sequentially
by the increment operation, thus being erroneously incremented when a power
failure occurs before updating the index \texttt{i}.

@@

Task-based models also suffer from data inconsistency when there is a WAR
dependence.  Prior systems tried to tackle the problem by a compiler-automated
redo-logging for the variables that is part of the dependence~\cite{alpaca}, or
by statically creating multiple copies of the problematic variable to ensure
that no task reads and writes the same copy~\cite{chain}.
%There are several run time strategies to ensuring task idempotence. One approach~\cite{alpaca}, is to identify non-volatile data involved in WAR dependences in a task (like DINO~\cite{dino} and Ratchet~\cite{ratchet} did for checkpoints), execute the task using private copies of those data, and commit the private copies on task completion. Another way is to statically create multiple versions of non-volatile data shared by tasks and ensure that no task reads and writes the same version~\cite{chain}. Regardless of the strategy, task-based systems execute statically-defined tasks atomically, completing in one or more attempt.

>>>>

Energy harvesting devices operate using energy extracted from sources such as
radio frequency transmissions and solar energy. These devices elide tethered
power or a battery, instead collecting energy into a capacitor, operating when
sufficient energy accumulates, and upon depleting the energy, turning off and
recharging.
There are several energy harvesting battery-less hardware platforms. For
instance, computational RFIDs---open-source TI MSP430-based~\cite{wolverine}
WISP~\cite{wisp5} (with its variants such as
WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or
NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones
such as~\cite{medusa_farsens_2017}. Other platforms include ambient backscatter
tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less
phone~\cite{talla_imwut_2017}.

@@

\begin{figure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-example.pdf}
        \caption{Subcaption}
        \label{fig:war-example}
    \end{subfigure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
        \caption{Subcaption}
        \label{fig:war-execution}
    \end{subfigure}
    \caption{Main caption.}
    \label{fig:war}
\end{figure}

Software running on an energy-harvesting device executes {\em intermittently}
because energy sources are not always available to harvest and buffer
sufficient operational energy. An intermittent execution is composed of
operating periods interspersed with power
failures~\cite{dino,chain,alpaca,ratchet}. The frequency of failures depends on
the size of the device's energy storage buffer: a larger buffer allows longer
operating periods.

A power failure clears volatile state (e.g., registers and SRAM) while
non-volatile memory (e.g., FRAM) persists. Upon a power failure, control flows
to a prior point in the execution: by default, to the beginning of {\tt
main()}. Early intermittent systems preserved progress by periodically
checkpointing volatile execution context to non-volatile
memory~\cite{mementos}, sometimes using hardware
support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}.

Checkpointing volatile state alone do not ensure data consistency when the
system can directly manipulate non-volatile memory~\cite{mspcdino}.
Especially, data can get inconsistent when code includes a
\emph{write-after-read} (WAR) dependence between operations that manipulate
non-volatile memory, because it allows {\em writes} from a failed attempt to
persist across failure and be read by the {\em
reads}~\cite{ratchet,dino,alpaca}\TODO{figure 2 dependent text ---see
Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become
inconsistent in an intermittent execution using the cyclic redundancy check
(CRC) code from MIBench2~\cite{hicks_mibench2_2016}.  Additional method to
maintain data consistent in the presence of WAR dependency is necessary for
correct execution.}

@@

Task-based models also suffer from data inconsistency when there is a WAR
dependence.  Prior systems tried to tackle the problem by a compiler-automated
redo-logging for the variables that is part of the dependence~\cite{alpaca}, or
by statically creating multiple copies of the problematic variable to ensure
that no task reads and writes the same copy~\cite{chain}.
